---
title: "BDI and RNA-seq Notebook"
output:
  html_document:
    df_print: paged
---

# Notebook for combining RNA-seq and BDI variables

## Load libraries

```{r load_libraries, message=FALSE, warning=FALSE}
library(knitr)
library(rgl)
library(pca3d)
knit_hooks$set(webgl = hook_webgl)

library(plyr)
library(tidyverse)
library(ggrepel)

# This copies the code provided by Vincent Q. Vu to avoid a dependance on devtools
# This file is licensed under the GPU GPL.
source("ggbiplot.r")

library(PRROC)
library(caret)
```

## Preprocess Data

Load and combine data for the **BDI** and **RNA-seq** variables. Only genes identified by **both** *edgeR* and *DEseq2* are used to build models later in this notebook. Genes identified using the decision tree bootstrapping method are included here as well.

```{r load_data, message = FALSE, warning=FALSE}
canine <- read_csv("canineCHOP81.csv")
mapping <- read_csv("../data/canineDLBCL_RNA-seqToOldSampleMapping.csv")
rna <- read_tsv("../data/counts_FPKM.tsv.gz")

# Read DE analysis, na.omit is to ensure both DESeq2 and EdgeR agree
resistant_v_sensitive_all <- read_tsv("resistant_vs_sensitive_all.csv") %>%
  select(Gene_ID, DE_in_Method, DESeq2_log2FC, EdgeR_log2FC, `Gene type`) %>%
  filter(`Gene type` == "protein_coding") %>%
  na.omit()

resistant_v_sensitive_pre <- read_tsv("resistant_vs_sensitive_pre.csv") %>%
  select(Gene_ID, DE_in_Method, DESeq2_log2FC, EdgeR_log2FC, `Gene type`) %>%
  filter(`Gene type` == "protein_coding") %>%
  na.omit()

# Flip the data and merge
rna_all <-
  rna %>%
  gather(sample, fpkm, -Gene_ID) %>%
  filter(Gene_ID %in% resistant_v_sensitive_all$Gene_ID) %>%
  spread(Gene_ID, fpkm) %>%
  merge(mapping, by = "sample") %>%
  as_tibble()

rna_pre <-
  rna %>%
  gather(sample, fpkm, -Gene_ID) %>%
  filter(Gene_ID %in% c(resistant_v_sensitive_pre$Gene_ID)) %>%
  spread(Gene_ID, fpkm) %>%
  merge(mapping, by = "sample") %>%
  as_tibble()

rna_combined <-
  rna %>%
  gather(sample, fpkm, -Gene_ID) %>%
  filter(Gene_ID %in% c(resistant_v_sensitive_pre$Gene_ID,
                        resistant_v_sensitive_all$Gene_ID)) %>%
  spread(Gene_ID, fpkm) %>%
  merge(mapping, by = "sample") %>%
  as_tibble()

# Cleanup
rna_all <-
  rna_all %>%
  filter(Condition == "Pre") %>%
  select(-sample, -Dog_Name, -Breed, -`BDI outcome`, -Condition) %>%
  mutate(Clinical_Outcome = factor(Clinical_Outcome)) %>%
  rename(ClinicalOutcome = Clinical_Outcome)

rna_pre <-
  rna_pre %>%
  select(-sample, -Dog_Name, -Breed, -`BDI outcome`) %>%
  rename(ClinicalOutcome = Clinical_Outcome)

rna_combined <-
  rna_combined %>%
  filter(Condition == "Pre") %>%
  select(-sample, -Dog_Name, -Breed, -`BDI outcome`, -Condition) %>%
  rename(ClinicalOutcome = Clinical_Outcome)

canine$X1 <- NULL
canine$BDIOutcme <- NULL

rm(rna)

```

## Variable information 

```{r variable_information}
bdi_auroc <-
  canine %>%
  summarise_if(is.numeric,
               ~ roc.curve(.[canine$ClinicalOutcome == "Sensitive"],
                           .[canine$ClinicalOutcome == "Resistant"])$auc) %>%
  gather %>% 
  arrange(desc(value))

bdi_auprc <-
  canine %>%
  summarise_if(is.numeric,
               ~ pr.curve(.[canine$ClinicalOutcome == "Sensitive"],
                          .[canine$ClinicalOutcome == "Resistant"])$auc.integral) %>%
  gather %>%
  arrange(desc(value))

rna_combined_auroc <-
  rna_combined %>%
  summarise_if(is.numeric,
               ~ roc.curve(.[canine$ClinicalOutcome == "Sensitive"],
                           .[canine$ClinicalOutcome == "Resistant"])$auc) %>%
  gather %>% 
  arrange(desc(value))

rna_combined_auprc <-
  rna_combined %>%
  summarise_if(is.numeric,
               ~ pr.curve(.[rna_combined$ClinicalOutcome == "Sensitive"],
                          .[rna_combined$ClinicalOutcome == "Resistant"])$auc.integral) %>%
  gather %>%
  arrange(desc(value))


```

### BDI variables AUROC

```{r bdi_auroc}
bdi_auroc
bdi_auroc %>% write_csv("tables/bdi_auroc.csv")
```

### BDI variables AUPR curve

```{r bdi_auprc}
bdi_auprc
bdi_auroc %>% write_csv("tables/bdi_auprc.csv")
```

### RNA variables ROC curve

```{r rna_auroc}
rna_combined_auroc
rna_combined_auroc %>% write_csv("tables/rna_auroc.csv")
```

### RNA variables PR curve

```{r rna_auprc}
rna_combined_auprc
rna_combined_auprc %>% write_csv("tables/rna_auprc.csv")
```

## Initial modelling attempt

Attempt to model the data using only the RNA-seq variables and BDI by themselves using regularized logistic regression and Leave-one-out cross validation and k-fold cross validation. These attempts do not yield any sets of hyperparameters which give a perfect classifier.

```{r initial_attempt_pure_variable_sets, message = FALSE, warning=FALSE}
loocv_training_helper <- function(data, method = "regLogistic", preP = c("center", "scale")) {
  train(
    factor(ClinicalOutcome) ~ .,
    data,
    trControl = trainControl("LOOCV"),
    preProcess = preP,
    method = method
  )
}

set.seed(12345)

rna_all_model <-
  loocv_training_helper(rna_all %>% select(-old_Sample_ID))

rna_pre_model <-
  loocv_training_helper(rna_pre %>% select(-old_Sample_ID))

rna_combined_model <-
  loocv_training_helper(rna_combined %>% select(-old_Sample_ID))

rna_combined_top_5 <-
  loocv_training_helper(rna_combined %>% select(ClinicalOutcome,
                                                ENSCAFG00000004237, ENSCAFG00000029984,
                                                ENSCAFG00000005330, ENSCAFG00000016518,
                                                ClinicalOutcome))

bdi_model <-
  loocv_training_helper(canine %>% select(-SampleName))

bdi_top03_model <- canine %>% select(LOF0_chop, ALLF1_pred, SDIP1_dox, ClinicalOutcome) %>% loocv_training_helper()
```

### Using only BDI variables

```{r bdi_only_print}
bdi_top03_model$results %>% arrange(desc(Kappa))
bdi_top03_model$results %>% write_csv("tables/bdi_best3_loocv.csv")

bdi_model$results %>% arrange(desc(Kappa))
bdi_model$results %>% write_csv("tables/bdi_loocv.csv")
```

### Using only RNA variables

```{r rna_only_print}
rna_combined_model$results %>% arrange(desc(Kappa))
rna_combined_model$results %>% write_csv("tables/rna_all_loocv.csv")

rna_combined_top_5$results %>% arrange((desc(Kappa)))
rna_combined_model$results %>% write_csv("tables/rna_top5_loocv.csv")
```

## Combine RNA and BDI

```{r bdi_rna_combine_models}

bdi_rna_combined <- merge(canine %>% select(-ClinicalOutcome),
                          rna_combined,
                          by.x = "SampleName", by.y = "old_Sample_ID")

set.seed(12345)

bdi_rna_combined %>%
  select(-SampleName) %>%
  loocv_training_helper() ->
  bdi_rna_all

set.seed(12345)

bdi_small_rna_all <-
  canine %>%
  select(LOF0_chop, ALLF1_pred, SDIP1_dox, SampleName) %>%
  merge(rna_combined, by.x = "SampleName", by.y = "old_Sample_ID") %>%
  select(-SampleName) %>%
  loocv_training_helper()

bdi_rna_select <-
  bdi_rna_combined %>%
  select(LOF0_chop, ALLF1_pred, SDIP1_dox,
        ENSCAFG00000004237, ENSCAFG00000029984,
        ENSCAFG00000005330, ENSCAFG00000016518,
        ClinicalOutcome, SampleName)

set.seed(12345)

bdi_rna_select_model <-
  bdi_rna_select %>%
  select(-SampleName) %>%
  loocv_training_helper()

set.seed(12345)

bdi_rna_select_no_preprocess <-
  bdi_rna_select %>%
  select(-SampleName) %>%
  loocv_training_helper(preP = NULL)

set.seed(12345)

rna_to_use <- c(
  "ENSCAFG00000004237", "ENSCAFG00000029984",
  "ENSCAFG00000005330", "ENSCAFG00000016518"
  )

pp <- list(
  center = rna_to_use,
  scale = rna_to_use
  )

bdi_rna_select_special_preprocess <-
  bdi_rna_select %>%
  select(-SampleName) %>%
  loocv_training_helper(preP = pp)

pp_2 <- list(
  center = rna_to_use,
  scale = c(rna_to_use,"ALLF1_pred", "SDIP1_dox", "LOF0_chop")
  )

bdi_rna_select_special_preprocess_2 <-
  bdi_rna_select %>%
  select(-SampleName) %>%
  loocv_training_helper(preP = pp_2)

```

### All BDI and all RNA variables

```{r all_combined}
bdi_rna_all$results
bdi_rna_all$results$Kappa %>% max()

bdi_small_rna_all$results
bdi_small_rna_all$results$Kappa %>% max()
```

### Selected BDI and RNA variables using the AUPRC metric

```{r select_combined}
bdi_rna_select_model$results
bdi_rna_select_no_preprocess$results
bdi_rna_select_special_preprocess$results
bdi_rna_select_special_preprocess_2$results

bdi_rna_select_model$results$Kappa %>% max
bdi_rna_select_no_preprocess$results$Kappa %>% max
bdi_rna_select_special_preprocess$results$Kappa %>% max
bdi_rna_select_special_preprocess_2$results$Kappa %>% max

bdi_rna_select_model$results %>% write_csv("tables/bdi_rna_loocv.csv")
bdi_rna_select_no_preprocess$results %>% write_csv("tables/bdi_rna_no_norm.csv")
bdi_rna_select_special_preprocess$results %>% write_csv("tables/bdi_rna_no_bdi_norm.csv")
bdi_rna_select_special_preprocess_2$results %>% write_csv("tables/bdi_rna_no_center.csv")

```

## Variable Analyis

```{r variable_analysis, fig.height=5, fig.width=6, fig.ext="svg", dev="svg"}
rna_all %>%
  ggplot(aes(ENSCAFG00000004237, fill = ClinicalOutcome)) +
  geom_dotplot(binwidth = 0.1, dotsize = 3) +
  geom_label_repel(aes(ENSCAFG00000004237, 0.1, label=old_Sample_ID), size = 3) +
  coord_cartesian(ylim = c(0, 0.2)) +
  cowplot::theme_cowplot()

rna_all %>%
  ggplot(aes(ENSCAFG00000016518, fill = ClinicalOutcome)) +
  geom_dotplot(binwidth = 0.1) +
  geom_label_repel(aes(ENSCAFG00000016518, 0.1, label=old_Sample_ID), size = 3) +
  coord_cartesian(ylim = c(0, 0.2)) +
  cowplot::theme_cowplot()

rna_all %>%
  ggplot(aes(ENSCAFG00000005330, fill = ClinicalOutcome)) +
  geom_dotplot(binwidth = 0.1, dotsize = 3) +
  geom_label_repel(aes(ENSCAFG00000005330, 0.1, label=old_Sample_ID), size = 3) +
  coord_cartesian(ylim = c(0, 0.2)) +
  cowplot::theme_cowplot()

rna_all %>%
  ggplot(aes(ENSCAFG00000029984, fill = ClinicalOutcome)) +
  geom_dotplot(binwidth = 0.1, dotsize = 3) +
  geom_label_repel(aes(ENSCAFG00000023923, 0.1, label=old_Sample_ID), size = 3) +
  coord_cartesian(ylim = c(0, 0.2)) +
  cowplot::theme_cowplot()

canine %>%
  ggplot(aes(LOF0_chop, fill = ClinicalOutcome)) +
  geom_dotplot(binwidth = 0.1, dotsize = 3) +
  geom_label_repel(aes(LOF0_chop, 0.1, label=SampleName), size = 3) +
  coord_cartesian(ylim = c(0, 0.2)) +
  cowplot::theme_cowplot()

canine %>%
  ggplot(aes(ALLF1_pred, fill = ClinicalOutcome)) +
  geom_dotplot(binwidth = 0.1) +
  geom_label_repel(aes(ALLF1_pred, 0.1, label=SampleName), size = 3) +
  coord_cartesian(ylim = c(0, 0.2)) +
  cowplot::theme_cowplot()

canine %>%
  ggplot(aes(SDIP1_dox, fill = ClinicalOutcome)) +
  geom_dotplot(binwidth = 0.1) +
  geom_label_repel(aes(SDIP1_dox, 0.1, label=SampleName), size = 3) +
  coord_cartesian(ylim = c(0, 0.2)) +
  cowplot::theme_cowplot()
```

## Leave one out testings (LOOT)

### LOOT function

```{r loot_function}
loot <- function(data, method = "regLogistic", preP = c("center", "scale")) {
  lapply(data$SampleName, function(X) {

    trn_vald_data <- data %>% filter(SampleName != X)

    train(
      factor(ClinicalOutcome) ~ .,
      method = method,
      trControl = trainControl("LOOCV"),
      data = trn_vald_data %>% select(-SampleName),
      preProcess = preP
    ) -> regLogiticModel
    
    pred = predict(regLogiticModel,
                     data %>% filter(SampleName == X)) %>% as.character()
    truth = (data %>% filter(SampleName == X))$ClinicalOutcome %>% as.character()
    
    self_score = sum(
      predict(
        regLogiticModel,
        data %>% filter(SampleName != X)
      ) ==
        (data %>% filter(SampleName != X))$ClinicalOutcome
    )
    
    validation_kappa <- regLogiticModel$results$Kappa %>% max()
    
    tibble(pred, truth, X, self_score, validation_kappa)
  }) %>% bind_rows()
}
```

### Run LOOT on all the data frames

```{r run_loot}

set.seed(12345)

bdi_s.loocv <-
  canine %>%
  loot()

set.seed(12345)

bdi_top_sloocv <-
  bdi_rna_select %>%
  select(ALLF1_pred, LOF0_chop, SDIP1_dox,
         SampleName, ClinicalOutcome) %>%
  loot()

set.seed(12345)

rna_combined_s.loocv <-
  rna_combined %>%
  rename(SampleName = old_Sample_ID) %>%
  loot()

set.seed(12345)

rna_subset_loot <-
  rna_combined %>%
  select(ENSCAFG00000004237, ENSCAFG00000029984,
         ENSCAFG00000005330, ENSCAFG00000016518,
         ClinicalOutcome, old_Sample_ID) %>%
  rename(SampleName = old_Sample_ID) %>%
  loot()

set.seed(12345)

bdi_rna_combined_s.loocv <-
  bdi_rna_select %>%
  loot()

set.seed(12345)

bdi_rna_combined_no_norm_s.loocv <-
  bdi_rna_select %>%
  loot(preP = NULL)

set.seed(12345)

bdi_rna_combined_ENSCAFG_norm_s.loocv <-
  bdi_rna_select %>%
  loot(preP = pp)

set.seed(12345)

bdi_rna_combined_special_norm_2.loot <-
  bdi_rna_select %>%
  loot(preP = pp_2)

```

### Print results

```{r print_s_loocv_results, warning=FALSE, message=FALSE}

summarize_result <- function(df, name) {
  my_table <- table(df$pred, df$truth)
  cm <- caret::confusionMatrix(my_table)
  
  res1 <- cm$byClass %>% enframe %>% rename(!!name := value)
  bind_rows(res1,
            cm$overall  %>% enframe %>% rename(!!name := value),
            tibble(name = "Average Validation Kappa", !!name := mean(df$validation_kappa)),
  )
}

all_loot_results <-
  bind_rows(
    bdi_s.loocv %>% mutate(input = "All BDI variables"),
    bdi_top_sloocv %>% mutate(input = "Top 3 BDI variables"),
    rna_combined_s.loocv %>% mutate(input = "All RNA variables"),
    rna_subset_loot %>% mutate(input = "Top 5 RNA variables"),
    bdi_rna_combined_s.loocv %>% mutate(input = "BDI and RNA variables w/ all normalization"),
    bdi_rna_combined_no_norm_s.loocv %>% mutate(input =  "BDI and RNA variables w/o normalization"),
    bdi_rna_combined_ENSCAFG_norm_s.loocv %>% mutate(input = "BDI and RNA variables w/ RNA normalization"),
    bdi_rna_combined_special_norm_2.loot %>% mutate(input = "BDI scaled and RNA variable w/ RNA normalization")
  ) %>% select('input', everything())

loot_summary <- 
  summarize_result(bdi_s.loocv, "All BDI variables") %>%
  inner_join(summarize_result(bdi_top_sloocv, "Top 3 BDI variables")) %>%
  inner_join(summarize_result(rna_combined_s.loocv, "All RNA variables")) %>%
  inner_join(summarize_result(rna_subset_loot, "Top 5 RNA variables")) %>%
  inner_join(summarize_result(bdi_rna_combined_s.loocv, "BDI and RNA variables w/ all normalization")) %>%
  inner_join(summarize_result(bdi_rna_combined_no_norm_s.loocv, "BDI and RNA variables w/o normalization")) %>%
  inner_join(summarize_result(bdi_rna_combined_ENSCAFG_norm_s.loocv, "BDI and RNA variables w/ RNA normalization")) %>%
  inner_join(summarize_result(bdi_rna_combined_special_norm_2.loot, "BDI scaled and RNA variable w/ RNA normalization"))

loot_summary %>%
  gather(key, value, -name) %>%
  spread(name, value) %>%
  arrange(Accuracy) %>%
  select(key, Accuracy, Precision, Recall, F1, Kappa, `Average Validation Kappa`) ->
  loot_summary

all_loot_results
loot_summary

all_loot_results %>% write_csv("tables/all_loot_results.csv")
loot_summary %>% write_csv("tables/all_loot_summary.csv")
```

## ENSEMBL to Gene name

Create an object to map the *ENSEMBL* genes to gene names if the file *annotations.csv* does not exist. This is done in a manner that prevents a reliance on the biomaRt package as it is not in CRAN.

```{r get_annotations, message = FALSE, warning=FALSE}
if(file.exists("annotations.csv")) {
  annotations = read_csv("annotations.csv")
  annotations$external_gene_name[is.na(annotations$external_gene_name)] = annotations$ensembl_gene_id[is.na(annotations$external_gene_name)]
} else {
  ensembl = biomaRt::useMart(
    biomart = "ENSEMBL_MART_ENSEMBL",
    dataset="cfamiliaris_gene_ensembl",
    host = 'www.ensembl.org',
    ensemblRedirect = FALSE
  )

  annotations = biomaRt::getBM(c("ensembl_gene_id", "external_gene_name"), mart = ensembl)

  # Remove this object so that the saved session does not depend on biomaRt
  rm(ensembl)
}
```

## Correlation plot

A correlation between **RNA-seq** and **BDI** variables is produced by the following function. 

```{r Correlation of BDI and RNAseq, fig.height=7, fig.width=6, fig.ext="svg", dev="svg"}

create_correlation_plot <- function(data, method = "spearman") {

  data %>%
    select(-ClinicalOutcome,
           -SampleName) %>%
    as.matrix() %>%
    cor(method = method) %>%
    as.data.frame() %>%
    rownames_to_column(var = 'var1') %>%
    gather(var2, value, -var1) %>%
    filter(grepl('ENSCAFG', var2),
           !grepl('ENSCAFG', var1)) %>%
    merge(annotations,
          by.x='var2',
          by.y = 'ensembl_gene_id',
          all.x = T) %>%
    mutate(external_gene_name = if_else(is.na(external_gene_name), var2, external_gene_name)) %>%
    mutate(external_gene_name = gsub("ENSCAFG000000", "E_", external_gene_name))->
    named_correlations

  named_correlations %>%
    dplyr::select(-var2) %>%
    spread('var1', value) -> corr.spread

  corr.spread %>%
    dplyr::select(-external_gene_name) %>%
    as.matrix() ->
    corr.matrix

  rownames(corr.matrix) <- corr.spread$external_gene_name

  corr.clust <- hclust(d = dist(x = corr.matrix)) %>% as.dendrogram() %>% order.dendrogram()
  corr.clust2 <- hclust(d = dist(x = corr.matrix %>% t)) %>% as.dendrogram() %>% order.dendrogram()

  named_correlations %>%
    mutate(external_gene_name = factor(external_gene_name,
                                     levels = corr.spread$external_gene_name[corr.clust])) %>%
    mutate(var1 = factor(var1,
                         levels = colnames(corr.matrix)[corr.clust2] )) %>%
    ggplot(aes(var1, external_gene_name, fill = value)) +
    geom_tile() +
    scale_fill_gradient2(limits = c(-1.0, 1.0), low = muted("blue"), high = muted("red")) +
    labs(x = "BDI Variable", y = "RNA-seq variable", fill = paste0(method,"\ncorrelation")) +
    cowplot::theme_cowplot() +
    theme(axis.text.x = element_text(angle = -90, hjust = 0, vjust = .5))
}

bdi_small_2 <-
  canine %>%
  gather(key, value, -SampleName, -ClinicalOutcome) %>%
  filter(key %in% c("SampleName", bdi_auprc$key[1:20])) %>%
  spread(key, value)

bdi_rna_combined_2 <- merge(bdi_small_2 %>% select(-ClinicalOutcome),
                          rna_combined,
                          by.x = "SampleName", by.y = "old_Sample_ID")


create_correlation_plot(bdi_rna_combined_2, "spearman")
create_correlation_plot(bdi_rna_combined_2, "pearson")
create_correlation_plot(bdi_rna_combined_2, "kendall")
```

```{r make_output_correlation_figures, message=FALSE}
pdf("figures/correlation.pdf", width = 6, height = 7)
create_correlation_plot(bdi_rna_combined_2, "pearson")
create_correlation_plot(bdi_rna_combined_2, "spearman")
create_correlation_plot(bdi_rna_combined_2, "kendall")
dev.off()
```

```{r cleanup_correlation_plots}
rm(bdi_small_2)
rm(bdi_rna_combined_2)
```

## PCA plots

```{r PCA_plots, fig.width=6, fig.height=7, fig.ext="svg", dev="svg"}
bdi_rna_select %>%
  select(-ClinicalOutcome, -SampleName) %>%
  prcomp(center = T, scale =T ) %>%
  ggbiplot(
    groups = bdi_rna_combined$ClinicalOutcome,
    labels = bdi_rna_combined$SampleName,
    ellipse = T
  ) +
  cowplot::theme_cowplot() +
  ggtitle("Selected variables, normalized")

bdi_rna_select %>%
  select(-ClinicalOutcome, -SampleName) %>%
  prcomp() %>%
  ggbiplot(
    groups = bdi_rna_combined$ClinicalOutcome,
    labels = bdi_rna_combined$SampleName,
    ellipse = T
  ) +
  cowplot::theme_cowplot() +
  ggtitle("Selected variables, not normalized")

bdi_rna_select %>%
  select(-ClinicalOutcome, -SampleName) %>%
  mutate(ENSCAFG00000004237 = scale(ENSCAFG00000004237),
         ENSCAFG00000029984 = scale(ENSCAFG00000029984),
         ENSCAFG00000005330 = scale(ENSCAFG00000005330),
         ENSCAFG00000016518 = scale(ENSCAFG00000016518),
         ALLF1_pred = ALLF1_pred,
         LOF0_chop = LOF0_chop,
         SDIP1_dox = SDIP1_dox) %>%
  prcomp() %>%
  ggbiplot(
    groups = bdi_rna_combined$ClinicalOutcome,
    labels = bdi_rna_combined$SampleName,
    ellipse = T
  ) +
  cowplot::theme_cowplot() +
  ggtitle("Selected variables, RNA normalized")

bdi_rna_select %>%
  select(-ClinicalOutcome, -SampleName) %>%
  mutate(ENSCAFG00000004237 = scale(ENSCAFG00000004237),
         ENSCAFG00000029984 = scale(ENSCAFG00000029984),
         ENSCAFG00000005330 = scale(ENSCAFG00000005330),
         ENSCAFG00000016518 = scale(ENSCAFG00000016518),
         ALLF1_pred = scale(ALLF1_pred, center = F),
         LOF0_chop = scale(LOF0_chop, center = F),
         SDIP1_dox = scale(SDIP1_dox, center = F)) %>%
  prcomp() %>%
  ggbiplot(
    groups = bdi_rna_combined$ClinicalOutcome,
    labels = bdi_rna_combined$SampleName,
    ellipse = T
  ) +
  cowplot::theme_cowplot() +
  ggtitle("Selected variables, no BDI centering")

bdi_rna_select %>%
  select(-ClinicalOutcome, -SampleName) %>%
  mutate(ENSCAFG00000004237 = scale(ENSCAFG00000004237),
         ENSCAFG00000029984 = scale(ENSCAFG00000029984),
         ENSCAFG00000005330 = scale(ENSCAFG00000005330),
         ENSCAFG00000016518 = scale(ENSCAFG00000016518),
         ALLF1_pred = scale(ALLF1_pred, scale = F),
         LOF0_chop = scale(LOF0_chop, scale = F),
         SDIP1_dox = scale(SDIP1_dox, scale = F)) %>%
  prcomp() %>%
  ggbiplot(
    groups = bdi_rna_combined$ClinicalOutcome,
    labels = bdi_rna_combined$SampleName,
    ellipse = T
  ) +
  cowplot::theme_cowplot() +
  ggtitle("Selected variables, no BDI scaling")

canine %>%
  select(-ClinicalOutcome, -SampleName) %>%
  prcomp(center = T, scale = T) %>%
  ggbiplot(
    groups = canine$ClinicalOutcome,
    labels = canine$SampleName
  ) +
  cowplot::theme_cowplot() +
  ggtitle("All BDI variables")

bdi_rna_select %>%
  select(ALLF1_pred, LOF0_chop, SDIP1_dox) %>%
  prcomp(center = T, scale = T) %>%
  ggbiplot(
    groups = canine$ClinicalOutcome,
    labels = canine$SampleName
  ) +
  cowplot::theme_cowplot() +
  ggtitle("Selected BDI variables")

bdi_rna_select %>%
  select(ENSCAFG00000004237, ENSCAFG00000005330,
         ENSCAFG00000029984, ENSCAFG00000016518) %>%
  prcomp(center = T, scale = T) %>%
  ggbiplot(
    groups = canine$ClinicalOutcome,
    labels = canine$SampleName
  ) +
  cowplot::theme_cowplot() +
  ggtitle("Selected RNA variables")

rna_combined %>%
  select(-ClinicalOutcome, -old_Sample_ID) %>%
  prcomp(center = T, scale = T) %>%
  ggbiplot(
    groups = rna_combined$ClinicalOutcome,
    labels = rna_combined$old_Sample_ID
  ) +
  cowplot::theme_cowplot()+
  ggtitle("All RNA variables")
```

### PCA in 3D

```{r PCA_3D, webgl=TRUE}
bdi_rna_select %>%
  select(-ClinicalOutcome, -SampleName) %>%
     prcomp(center = T, scale =T ) %>%
     pca3d::pca3d(group = bdi_rna_combined$ClinicalOutcome)
```

# Save Everything out

```{r save_everything_out}
save.image(file="knitr.rda")

bdi_rna_select_special_preprocess$finalModel$W %>%
  as_tibble %>%
  write_csv("tables/model_weights.csv")
```
